Legacy Data Confounds Genomics Studies

### 2019-08-12 LatexManuscript/MBE/legacy.tex
commit while modifying things

In particular, the latter two studies  identified a heterogeneous mutational signature within 1kGP Japanese individuals.
This heterogeneity is intriguing because differences in germline signatures accumulate over many generations.
A systematic difference within the Japanese population would suggest sustained environmental or genetic differences across sub-populations within Japan with little to no gene flow.
\todo{This signature had been partially reproduced at continental resolution in the SGDP. 
However, due to smaller sample sizes, statistical power lacked to confirm the signature.
Moreover, the result had yet to be reproduced in a Japanese dataset, so we decided to follow up on this observation, by using a newly sequenced dataset of Japanese individuals from Nagahama \citep{Harris2017a,Aikens2018}.}
While we were unable to reproduce the mutational heterogeneity within the Japanese population, we could trace back the source of the discrepancy to a technical artefact in the 1kGP data.
In addition to creating biases in mutational signatures, this artefact leads to spurious imputation results which have found their way in recent publications and online resources.

@@

$Q$-associated variants are distributed across the genome, with chromosome 1 showing an excess of such variants, and other chromosomes being relatively uniform (Figure \ref{CountPerChrom}).
\todo{Chromosome 1 shows enrichments in the  *AC${\rightarrow}$*CC signal compared to other chromosomes despite accounting for chromosomal length (Figure \ref{MutEnrichment_chrom}).} 
At a 10kb scale, we also see rather uniform distribution with a small number of regions showing an enrichment for such variants (Figure \ref{VariantDensity}).
An outlying 10kb region in chromosome 17 (bases : 22,020,000 to  22,030,000) has 35 $Q$-associated variants.
Distribution of association statistics in this region is provided in Figure \ref{Top10}.
By contrast, variants that do not pass the 1kGP strict mask are more unevenly distributed across the genome (Figure \ref{VariantDensity}).

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=10cm,keepaspectratio]{mutational_enrichment_per_chrom.jpg}
\caption{Mutation spectrum enrichment of  $Q$-associated variants across each chromosome.
Chromosome 1 has a strong enrichment in *AC${\rightarrow}$*CC mutational pattern.}  
\label{MutEnrichment_chrom}
\end{figure}

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=11cm,keepaspectratio]{CountPerChrom.jpg}
\caption{The proportion of $Q$-associated variants per chromosome.}  
\label{CountPerChrom}
\end{figure}

>>>>

In particular, the latter two studies  identified a heterogeneous mutational signature within 1kGP Japanese individuals.
This heterogeneity is intriguing because differences in germline signatures accumulate over many generations.
A systematic difference within the Japanese population would suggest sustained environmental or genetic differences across sub-populations within Japan with little to no gene flow.

\todo{
An excess of this signature has been observed in East Asians compared to Afriacas in the SGDP. 
At a continental resolution, this signature has been partially reproduced in the SGDP -- an independent dataset with smaller sample size from the 1kGP \citep{Aikens2018, Harris2017a}. 
Nevertheless, the heterogeneous signature reported by \cite{Harris2017a} had yet to be reproduced in a Japanese dataset, so we decided to follow up on this observation, by using a newly sequenced dataset of Japanese individuals from Nagahama.}
The herertogeneous signatrue reported by harris et al had not , so we decided.
While we were unable to reproduce the mutational heterogeneity within the Japanese population, we could trace back the source of the discrepancy to a technical artefact in the 1kGP data.
In addition to creating biases in mutational signatures, this artefact leads to spurious imputation results which have found their way in recent publications and online resources.

@@

$Q$-associated variants are distributed across the genome, with chromosome 1 showing an excess of such variants, and other chromosomes being relatively uniform (Figure \ref{CountPerChrom}A).
\todo{Chromosome 1 shows strong enrichments in the  *AC${\rightarrow}$*CC signal compared to other chromosomes (Figure \ref{CountPerChrom}B).} 
At a 10kb scale, we also see rather uniform distribution with a small number of regions showing an enrichment for such variants (Figure \ref{VariantDensity}).
An outlying 10kb region in chromosome 17 (bases : 22,020,000 to  22,030,000) has 35 $Q$-associated variants.
Distribution of association statistics in this region is provided in Figure \ref{Top10}.
By contrast, variants that do not pass the 1kGP strict mask are more unevenly distributed across the genome (Figure \ref{VariantDensity}).

@@

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=11cm,keepaspectratio]{CountPerChrom.jpg}
\caption{
\textbf{A}
The proportion of $Q$-associated variants per chromosome.
\textbf{B}
Mutation spectrum enrichment of  $Q$-associated variants across each chromosome.
Chromosome 1 has a strong enrichment in *AC${\rightarrow}$*CC mutational pattern.
}  
\label{CountPerChrom}
\end{figure}
